---
title: "Multivariate, muliple-outcome mixed effects model"
author: "TJ Mahr"
date: "April 3, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE)
```

```{r}
library(dplyr)
library(readr)
d <- read_csv("./data/NIHMS557790-supplement-Dataset_long_format.csv")
d$outcome <- ifelse(d$var == 1, "depression", "quality")
d
```

```{r}
library(ggplot2)
ggplot(d) + 
  aes(x = time, y = y, color = factor(tx)) + 
  geom_point(position = position_jitterdodge(dodge.width = .3, jitter.width = .1), alpha = .25) + 
  stat_summary(position = position_dodge(width = .3), size = 1) + 
  facet_grid(var ~ ., labeller = label_both)

ggplot(d) + 
  aes(x = time, y = y, color = factor(tx)) + 
  geom_line(aes(group = pid)) + 
  facet_grid(var ~ ., labeller = label_both)
```


## Univariate outcome model

Let's consider a single outcome.

```{r}
dl <- d %>%
  select(pid, time, tx, var, y) %>%
  mutate(outcome = paste0("var", var)) %>%
  select(pid, time, tx, outcome, value = y)
dl_1 <- dl %>% filter(outcome == "var1")
dl_1
```

For one outcome, it's a varying slope model. They give this equation.

$$y_{1ij}=β_{10}+β_{11}Time_{ij}+β_{12}Tx_j+β_{13}Time_{ij}Tx_j+u_{1j}+v_{1j}Time_{ij}+e_{1ij}$$

And describe it like so. Notes in brackets are my addition.

> Focusing on Equation (1), y1ij [`value`] is the depression outcome at time i for person
j. β10 is the overall intercept [`~ 1`], and like all intercepts, it represents the
expected depression value (i.e., mean) of the outcome when all predictors are
equal to 0, in this case when Timeij = 0 (Baseline) and Txj = 0 (control). [...] 
β11 is the average rate of change in depression symptoms
during treatment for the control condition [`~ time`], β12 is the mean difference between
CBT and control at baseline[`~ tx`], β13 is the difference in rate of change between CBT
and control (i.e., the treatment effect) [`~ tx:time`], u1j is a random effect representing
person-specific differences at baseline (i.e., unique baseline values for each
participant) [`~ (1 | pid)`], v1j is a random effect representing person-specific differences in
change during treatment (i.e., unique rate of change for each participant) [`~ (time | pid)`], and
e1ij is residual error.

```{r}
library(lme4)
arm::display(lmer(value ~ time * tx + (1 + time | pid), dl_1))
```

This matches the top left corner of the table here: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4119868/table/T2/

```{r}

```


## Multivariate outcomes model

It seems like it's harder to write the independent outcomes model than it is
write the related outcomes model. (The matrix in (7) looks harder to code.) So I
will do the full, related outcomes model.

```{r}
help
```



First, let's ignore nesting of data within participants and try a "seemingly
unrelated regression" model. I copied this from the Stan manual.

```{r}
sur_stan_code <- "
data {
  int<lower=0> N; // num observations
  int<lower=1> J; // num predictors
  int<lower=1> K; // num outcomes 
  vector[J] x[N]; // predictor matrix
  vector[K] y[N]; // outcome matrix
}
parameters {
  matrix[K, J] beta;
  corr_matrix[K] corr_outcomes;
  vector<lower=0>[K] scale_outcomes;
}
model {
  vector[K] mu[N];
  
  corr_outcomes ~ lkj_corr(2);
  scale_outcomes ~ cauchy(0, 2.5);

  for (n in 1:N) {
    mu[n] = beta * x[n];
  }

  y ~ multi_normal(mu, quad_form_diag(corr_outcomes, scale_outcomes));
}
"
```

```{r, cache=TRUE, cache.vars = "sur_stan_code"}
sur_model <- stan_model(model_code = sur_stan_code)
```


```{r}
library(rstan)
dwide <- d %>% 
  select(pid:y, outcome) %>% 
  tidyr::spread(outcome, y) %>% 
  print()

sur_data <- list(
  N = nrow(dwide),
  J = ncol(model.matrix(~ time * tx, dwide)),
  K = ncol(model.matrix(~ 0 + depression + quality, dwide)),
  x = model.matrix(~ time * tx, dwide),
  y = model.matrix(~ 0 + depression + quality, dwide)
)
```

```{r, cache=TRUE, cache.vars = "sur_model"}
sur_results <- sampling(sur_model, sur_data)
sur_results
```


Now, we should add some infrastructure for random effects.

```{r}
rsur_stan_code <- "
data {
  int<lower=0> N; // num observations
  int<lower=1> J; // num predictors
  int<lower=1> K; // num outcomes 
  
  int<lower=1> P; // num participants 
  int<lower=1> L; // num participant-level predictors 
  
  row_vector[L] u[P]; // participant-level predictors (interact with randomly varying predictors)
  vector[J] x[N]; // randomly varying predictors matrix
  vector[K] y[N]; // outcome matrix
}
parameters {
  matrix[K, J] beta;  
  matrix[L, K * J] gamma; 

  matrix[K * J, K * J] ranefs;
  corr_matrix[K * J] corr_ranef;
  vector<lower=0>[K * J] scale_ranef;

  corr_matrix[K] corr_outcomes;
  vector<lower=0>[K] scale_outcomes;
}
model {
  vector[K] mu[N];

  corr_ranef ~ lkj_corr(2);
  scale_ranef ~ cauchy(0, 2.5);
  
  corr_outcomes ~ lkj_corr(2);
  scale_outcomes ~ cauchy(0, 2.5);

  to_vector(gamma) ~ normal(0, 5);
  {
    row_vector[J] u_gamma[P];
    for (p in 1:P) {
      u_gamma[p] = u[p] * gamma;
    }
    to_vector(beta) ~ multi_normal(u_gamma, quad_form_diag(corr_ranef, scale_ranef));
  }

  for (n in 1:N) {
    mu[n] = beta * x[n];
  }

  y ~ multi_normal(mu, quad_form_diag(corr_outcomes, scale_outcomes));
}
"
```



```{r}
omg <- stan_model(model_code = rsur_stan_code)
```

```{r}
d_treat <- dwide %>% select(pid, tx) %>% unique
sur_data <- list(
  N = nrow(dwide),
  J = ncol(model.matrix(~ time, dwide)),
  K = ncol(model.matrix(~ 0 + depression + quality, dwide)),
  P = length(unique(dwide$pid)),
  L = ncol(model.matrix(~ 1 + tx, dwide)),
  u = model.matrix(~ 1 + tx, d_treat),
  x = model.matrix(~ time, dwide),
  y = model.matrix(~ 0 + depression + quality, dwide)
)

 
#   int<lower=0> N; // num observations
#   int<lower=1> J; // num predictors
#   int<lower=1> K; // num outcomes 
#   
#   int<lower=1> P; // num participants 
#   int<lower=1> L; // num participant-level predictors 
#   
#   row_vector[L] u[P]; // participant-level predictors (interact with randomly varying predictors)
#   vector[J] x[N]; // randomly varying predictors matrix
#   vector[K] y[N]; // outcome matrix
sur_results <- sampling(omg, sur_data)
sur_results
```

Hmmm... I still haven't cracked this one yet.

```
Rejecting initial value:
  Error evaluating the log probability at the initial value.

Rejecting initial value:
  Error evaluating the log probability at the initial value.

Rejecting initial value:
  Error evaluating the log probability at the initial value.

Rejecting initial value:
  Error evaluating the log probability at the initial value.

Initialization between (-2, 2) failed after 100 attempts. 
 Try specifying initial values, reducing ranges of constrained values, or reparameterizing the model.
[1] "Error in sampler$call_sampler(args_list[[i]]) : Initialization failed."
[1] "error occurred during calling the sampler; sampling not done"
```
