---
title: "Multivariate, muliple-outcome mixed effects model"
author: "TJ Mahr"
date: "May 18, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE)
```

I've been trying to fit a multivariate mixed effects model, as described in
[this article](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4119868/) using pure
Stan code. The article has dummy example of this kind of model that I am using
to validate candidate models.

```{r, message = FALSE}
library(dplyr)
library(readr)
library(brms)
library(rstan)
options(mc.cores = parallel::detectCores())

d <- read_csv("./data/NIHMS557790-supplement-Dataset_long_format.csv")
dwide <- d %>% 
  mutate(outcome = ifelse(var == 1, "depression", "quality")) %>% 
  select(pid:y, outcome) %>% 
  tidyr::spread(outcome, y)
dwide
```

But it turns out to be trivial in brms.

Fit the model:

```{r, eval = FALSE}
model <- brm(
  cbind(depression, quality) ~ time * tx + (time | 2 | pid), 
  data = dwide, iter = 10000, 
  save_dso = TRUE, 
  save_model = "model.stan")
```

```{r, echo = FALSE}
load("model.Rdata")
```

Here's the fit.

```{r}
model
```

Load up the results from the table in the paper.

```{r}
paper <- list(
  "depression_Intercept" = −.08,
  "depression_time"	= .06, 
  "depression_tx"	=	.16,
  "depression_time:tx" = −.41,
  
  "quality_Intercept"	= −.08,
  "quality_time"	= .002,
  "quality_tx"	= .08,
  "quality_time:tx" = −.09,
  
  "sd(depression_Intercept)" = sqrt(.49),
  "sd(depression_time)" = sqrt(.04),
  # "cov(depression_Intercept,depression_time)" = .004,
  "cor(depression_Intercept,depression_time)" = .004 / (sqrt(.49) * sqrt(.04)),
  
  "sd(quality_Intercept)" =	sqrt(.73),
  "sd(quality_time)" =	sqrt(.14),
  # "cov(quality_Intercept,quality_time)" = −.14,
  "cor(quality_Intercept,quality_time)" = −.14 / (sqrt(.73) * sqrt(.14)),
  
  # "cov(depression_Intercept,quality_Intercept)"	= .16,
  "cor(depression_Intercept,quality_Intercept)"	= .16 / (sqrt(.49) * sqrt(.73)),
  # "cov(depression_Intercept,quality_time)" =	−.04,
  "cor(depression_Intercept,quality_time)" =	−.04 / (sqrt(.49) * sqrt(.14)),
  # "cov(depression_time,quality_Intercept)" = −.04,
  "cor(depression_time,quality_Intercept)" = −.04 / (sqrt(.04) * sqrt(.73)),
  # "cov(depression_time,quality_time)"	= .06,
  "cor(depression_time,quality_time)"	= .06 / (sqrt(.04) * sqrt(.14)),
  
  "sigma(depression)" =	sqrt(.39),
  "sigma(quality)" = sqrt(.30),
  # "cov(depression,quality)" =	.15,
  "rescor(depression,quality)" =	.15 / (sqrt(.39) * sqrt(.30))
)
```

Some junk to clean up the column names

```{r}
df <- as.data.frame(model)
names(df) <- names(as.data.frame(model)) %>% 
  stringr::str_replace("time.tx", "time:tx") %>% 
  stringr::str_replace("^b_", "") %>% 
  stringr::str_replace("^sd(.+)$", "sd(\\1)") %>% 
  stringr::str_replace("^cor(.+)$", "cor(\\1)") %>% 
  stringr::str_replace("^rescor(.+)$", "rescor(\\1)") %>% 
  stringr::str_replace("^sigma_(.+)$", "sigma(\\1)") %>% 
  stringr::str_replace("pid", "") %>% 
  stringr::str_replace("[(]_+", "(") %>% 
  stringr::str_replace("__+", ",")
df <- df %>% select(1:21)  %>% tibble::as.tibble()
```

Compare the MLE and MCMC estimates!

```{r}
ordering <- names(df)
dfl <- df %>% tidyr::gather(parameter, value)
paper_df <- tibble::enframe(paper, name = "parameter") %>% 
  tidyr::unnest(value)

library(ggplot2)
ggplot(dfl) + 
  aes(x = factor(parameter, levels = rev(ordering)), y = value) + 
  stat_summary(aes(color = "brms fit"), fun.data = median_hilow,  
               fun.args = list(conf.int = .95)) + 
  geom_point(aes(color = "article"), data = paper_df) + 
  coord_flip() + 
  scale_color_manual(values  = c("red", "black")) + 
  labs(y = NULL, x = NULL, colour = "source") + 
  theme(legend.position = "bottom")
```


